A Multiple-Set Overlapping-Domain Decomposed Monte Carlo Synthetic Acceleration Method for Linear Systems

### 2013-09-06 doc/sna2013/paper/slattery_evans_wilson.tex
fixing relative weight cutoff definition in dissertation

\abstract{We present a multiple-set overlapping-domain decomposed
  strategy for parallelizing the Monte Carlo Synthetic Acceleration
  (MCSA) method. MCSA methods use the Neumann-Ulam class of Monte
  Carlo solvers for linear systems to accelerate the fixed point
  method on which the Neumann-Ulam solvers are based. To effectively
  parallelize MCSA methods requires the parallelization of the
  underlying Neumann-Ulam solvers. To do this in a domain decomposed
  environment, we borrow strategies traditionally implemented in Monte
  Carlo particle transport to parallelize the problem. The
  multiple-set overlapping-domain decomposition algorithm is presented
  along with parallel scaling data for the resulting implementation
  using the Titan Cray XK7 machine at Oak Ridge National Laboratory.}

@@

For some time, the particle transport community has utilized Monte
Carlo methods for the solution of transport problems
\cite{lewis_computational_1993}. The partial differential equation
(PDE) community has focused on various deterministic methods for
solutions to linear problems \cite{saad_iterative_2003,
  kelley_iterative_1995}. In between these two areas are a not widely
known group of Monte Carlo methods for solving sparse linear systems
\cite{forsythe_matrix_1950, hammersley_monte_1964,
  halton_sequential_1962, halton_sequential_1994}. In recent years,
these methods have been further developed for radiation transport
problems in the form of Monte Carlo Synthetic Acceleration (MCSA)
\cite{evans_monte_2009, evans_monte_2012} but have yet to be applied
to more general sparse linear systems commonly generated by the
computational physics community. Compared to other methods in this
regime, MCSA offers three attractive qualities; (1) the physics
operator need not be symmetric or positive-definite, (2) the
stochastic nature of the solution method provides a natural solution
to some aspects of the issue of resiliency, and (3) is amenable to
parallelization using modern methods developed by the transport
community \cite{wagner_hybrid_2010}. The development a parallel MCSA
algorithm using techniques is the primary contribution of this work.

As leadership class machines move beyond the petascale, new algorithms
must be developed that leverage their strengths and adapt to their
shortcomings. Basic research is required now to advance methods in
time for these new machines to become operational. As machines begin
to operate at hundreds of petaflops peak performance and beyond,
trends toward reduced energy consumption will require incredibly high
levels of concurrency to achieve the desired computation rates
\cite{kogge_using_2011}. The end result of these hardware changes is
that the larger number of low-powered processors will be prone to both
soft failures such as bit errors in floating point operations and hard
failures where the data owned by that processor cannot be
recovered. Because these failures are predicted to be common,
resilient solver technologies are required to overcome these events at
the application level. With linear solvers based on Monte Carlo
techniques, such issues are potentially alleviated by statistical
arguments. In the case of soft failures, isolated floating point
errors in the Monte Carlo simulation are absorbed within tally
statistics. Completely losing memory during a hard failure is treated
as a high variance event where some portion of the Monte Carlo
histories and subsequently the solution are lost with other histories
maintained by replicating the problem and combining the results
through superposition.

In this paper, we will present the Monte Carlo methods of Neumann and
Ulam and then outline Monte Carlo Synthetic Acceleration
methods. Next, we present a parallel algorithm for MCSA based on the
work of Brunner and Brantley \cite{brunner_efficient_2009} and Wagner
et. al. \cite{wagner_hybrid_2010}. We then perform a series of
parallel scaling studies to characterize the parallel
algorithm. Finally, we summarize the work and offer future work that
may improve the algorithm.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Monte Carlo Methods}
Monte Carlo Synthetic Acceleration (MCSA) methods rely on the
Neumann-Ulam class of linear solvers to accelerate a stationary
method. These Monte Carlo methods work by sampling a distribution with
an expectation value equivalent to that of the inverted
operator. Neumann and Ulam's work was first published in 1950 by
Forsythe and Leibler \cite{forsythe_matrix_1950}. In the following
decade, the books of Hammersley and Handscomb
\cite{hammersley_monte_1964} and Spanier and Gelbard
\cite{spanier_monte_1969} present additional detail on these solvers.
In this section we briefly present the Neumann-Ulam Monte Carlo method
for solving linear systems and then present the Monte Carlo Synthetic
Acceleration method.

\subsection{Neumann-Ulam Method}
We seek solutions of the general linear problem in the following form:
\begin{equation}
  \mathbf{A} \mathbf{x} = \mathbf{b}\:,
  \label{eq:linear_problem}
\end{equation}
where $\mathbf{A} \in \mathbb{R}^{N \times N}$ is a matrix operator
such that $\mathbf{A} : \mathbb{R}^{N} \rightarrow \mathbb{R}^{N}$,
$\mathbf{x} \in \mathbb{R}^N$ is the solution vector, and $\mathbf{b}
\in \mathbb{R}^N$ is the forcing term.  For a given linear operator
$\mathbf{A}$, we can use diagonal splitting to define the
\textit{iteration matrix}, $\mathbf{H}$:
\begin{equation}
  \mathbf{H} = \mathbf{I} - \mathbf{A}\:,
  \label{eq:linear_mc_iteration_matrix}
\end{equation}
such that we are solving the system:
\begin{equation}
  \mathbf{x} = \mathbf{H} \mathbf{x} + \mathbf{b}\:.
  \label{eq:richardson_split}
\end{equation}
We can then form an alternative representation for $\mathbf{A}^{-1}$
by generating the \textit{Neumann series} solution:
\begin{equation}
  \mathbf{A}^{-1}\mathbf{b} = \sum_{k=0}^{\infty}
  \mathbf{H}^k\mathbf{b} = \mathbf{x}\:,
  \label{eq:neumann_solution}
\end{equation}
which will converge if the spectral radius of $\mathbf{H}$ is less
than 1. If we expand the summation with a succession of matrix-vector
multiply operations, we arrive at an alternative perspective of this
summation by considering the $i^{th}$ component of the solution
vector:
\begin{equation}
  x_i = \sum_{k=0}^{\infty}\sum_{i_1}^{N}\sum_{i_2}^{N}\ldots
  \sum_{i_k}^{N}h_{i,i_1}h_{i_1,i_2}\ldots h_{i_{k-1},i_k}b_{i_k}\:,
  \label{eq:expanded_neumann_solution}
\end{equation}
which can interpreted as a series of transitions between states,
\begin{equation}
 \nu = i \rightarrow i_1 \rightarrow \cdots \rightarrow i_{k-1}
 \rightarrow i_{k}\:,
  \label{eq:mc_walk_permutation}
\end{equation}
in $\mathbf{H}$ where $\nu$ is interpreted as a particular sequence
permutation. We can generate these sequences of transitions through
Monte Carlo random walks by assigning them both a probability and
weight. Using the adjoint form \cite{spanier_monte_1969},
probabilities are column-scaled:
\begin{equation}
  p_{ij} = \frac{|h_{ji}|}{\sum_j |h_{ji}|}\:,
  \label{eq:adjoint_probability}
\end{equation}
such that we expect to select a new state, $j$, from the current state
in the random walk, $i$, by sampling by sampling the probability
distribution function formed by the $i^{th}$ of the probability
matrix. The transition weight is then defined as:
\begin{equation}
  w_{ij} = \frac{h_{ji}}{p_{ij}}\:.
  \label{eq:adjoint_weight}
\end{equation}
The initial state $i_0$ of the random walk is determined by sampling
the source vector $\mathbf{b}$ with probabilities:
\begin{equation}
  P_{(i_0=i)}(\nu) = \frac{|b_i|}{||\mathbf{b}||_1}\:,
  \label{eq:adjoint_source_probability}
\end{equation}
with a random walk starting weight of:
\begin{equation}
  W_0 = ||\mathbf{b}||_1 \frac{b_i}{|b_i|}\:.
  \label{eq:adjoint_starting_weight}
\end{equation}
The contribution to the solution in state $j$ from a particular random
walk permutation of $k$ events is then the \textit{collision
  estimator}:
\begin{equation}
  X_{j}(\nu) = \sum_{m=0}^k W_{m} \delta_{i_m,j}\:,
  \label{eq:adjoint_permutation_contribution}
\end{equation}
where the Kronecker delta indicates that the tally contributes only in
the current state, $i_m$, of the random walk and 
\begin{equation}
  W_{m} = W_0 w_{i_0,i_1} w_{i_1,i_2} \cdots w_{i_{m-1},i_m}\:.
  \label{eq:adjoint_permutation_weight}
\end{equation}
Finally, the expectation value using all permutations is:
\begin{equation}
  E\{X_j\} = \sum_{\nu} P_{\nu} X_{j}(\nu)\:,
  \label{eq:adjoint_expectation_value}
\end{equation}
with
\begin{equation}
  P_{\nu} = P_{(i_0=i)} p_{i,i_1} p_{i_1,i_2} \cdots p_{i_{k-1},i_k}\:.
  \label{eq:adjoint_permutation_probability}
\end{equation}

\subsection{Monte Carlo Synthetic Acceleration}
Using the ideas of Halton's residual method
\cite{halton_sequential_1994}, Evans and Mosher recently developed a
Monte Carlo solution method that was not prohibited severely by the
quality of the initial guess for the system \cite{evans_monte_2009}
and later applied it more rigorously as a solution mechanism for the
radiation diffusion equation \cite{evans_monte_2012}. With their new
methods, they achieved identical numerical results as and marginally
better performance than conventional subspace methods. Their approach
was instead to use residual Monte Carlo as a synthetic acceleration
for a stationary method.

The \textit{Fixed-Point Monte Carlo Synthetic Acceleration} (MCSA)
method is defined as:
\begin{subequations}
  \begin{gather}
    \label{eq:mcsa_1}
    \mathbf{r}^{k} = \mathbf{b} - \mathbf{A}\mathbf{x}^{k}\:,\\
    \label{eq:mcsa_2}
    \mathbf{x}^{k+1/2} = \mathbf{x}^k + \mathbf{r}^k\:,\\
    \label{eq:mcsa_3}
    \mathbf{r}^{k+1/2} = \mathbf{b} - \mathbf{A}\mathbf{x}^{k+1/2}\:,\\
    \label{eq:mcsa_4}
    \mathbf{A}\delta\mathbf{x}^{k+1/2} = \mathbf{r}^{k+1/2}\:,\\
    \label{eq:mcsa_5}
    \mathbf{x}^{k+1} = \mathbf{x}^{k+1/2} + \delta \mathbf{x}^{k+1/2}\:,
  \end{gather}
  \label{eq:mcsa}
\end{subequations}
where $\mathbf{r}$ is the residual vector, $\delta\mathbf{x}$ is the
Monte Carlo correction vector, and $k$ indicates the iteration
index. Here, a Neumann-Ulam Monte Carlo method is used to generate the
solution correction in Eq~\ref{eq:mcsa_4} from the residual and
Richardson's iteration in the first step has been rewritten as a
residual correction. Using Monte Carlo in this way achieves the same
effect as Halton's method, decoupling its convergence rate from the
overall convergence rate of the method. Here, the approximate Monte
Carlo solution is not driven to a particular convergence as it merely
supplies a correction for the initial guess generated by Richardson's
iteration. Rather, only a set number of random walk permutations are
required using the Neumann-Ulam method to generate the correction. In
addition to the Monte Carlo solver parameters dictating the number of
histories and weight cutoff, the outer MCSA iterations also have the
following stopping criteria:
\begin{equation}
  ||\mathbf{r}||_\infty < \epsilon \ ||\mathbf{b}||_\infty\:,
  \label{eq:mcsa_stopping_criteria}
\end{equation}
where $\epsilon$ is a user-defined parameter. We refer the reader to
\cite{evans_monte_2012} and \cite{evans_monte_2009} for a more
detailed explanation and analysis of MCSA and the Neumann-Ulam
methods.

@@

\subsection{Overlapping Domain Decomposition}
\label{subsec:overlapping_domain_decomp}
In the next set of scaling studies, we will explore how adding overlap
to the subdomains in the problem affects run times and parallel
performance as compared to the case with pure domain decomposition. To
do this, we utilize the set of analytic relations developed in
Appendix~\ref{chap:parallel_theory} to design these numerical
experiments. For the neutron diffusion problem used in these scaling
studies, the linear operator had a spectral radius of 0.787 and a
weight cutoff of \sn{1}{-2} was used with the Neumann-Ulam
solver. Using Eq~(\ref{eq:analytic_k}) gives an expected random walk
length of 19.22 transitions for any given history in the
problem. Using the mean-chord approximation given by
Eq~(\ref{eq:mean_chord_domain_leakage}), we then expect approximately
5.3\% of the histories born in a particular domain to leak out after
the first transport stage. By adding overlap, we aim to reduce this
fraction of communicated histories and therefore boost the parallel
performance of the algorithm.

>>>>

\abstract{We present a multiple-set overlapping-domain decomposed
  strategy for parallelizing the Monte Carlo Synthetic Acceleration
  method. Monte Carlo Synthetic Acceleration methods use the
  Neumann-Ulam class of Monte Carlo solvers for linear systems to
  accelerate a fixed-point iteration sequence. Effective parallel
  algorithms for these methods require the parallelization of the
  underlying Neumann-Ulam solvers. To do this in a domain decomposed
  environment, we borrow strategies traditionally implemented in Monte
  Carlo particle transport to parallelize the problem. The parallel
  Neumann-Ulam and multiple-set overlapping-domain decomposition
  algorithm is presented along with parallel scaling data for the
  resulting implementation using the Titan Cray XK7 machine at Oak
  Ridge National Laboratory.}

@@

For some time, the particle transport community has utilized Monte
Carlo methods for the solution of transport problems
\cite{lewis_computational_1993}. The partial differential equation
(PDE) community has focused on various deterministic methods for
solutions to linear problems \cite{saad_iterative_2003,
  kelley_iterative_1995}. In between these two areas are a not widely
known group of Monte Carlo methods for solving sparse linear systems
\cite{forsythe_matrix_1950, hammersley_monte_1964,
  halton_sequential_1962, halton_sequential_1994}. In recent years,
these methods have been further developed for radiation transport
problems in the form of Monte Carlo Synthetic Acceleration (MCSA)
\cite{evans_monte_2009, evans_monte_2012} but have yet to be applied
to more general sparse linear systems commonly generated by the
computational physics community. Compared to other methods in this
regime, MCSA offers three attractive qualities; (1) the physics
operator need not be symmetric or positive-definite, (2) the
stochastic nature of the solution method potentially provides a
natural solution to some aspects of the issue of resiliency, and (3)
is amenable to parallelization using modern methods developed by the
transport community \cite{wagner_hybrid_2010}. The development of a
parallel MCSA algorithm using these techniques is the primary
contribution of this work.

As leadership class machines move beyond the petascale, new algorithms
must be developed that leverage their strengths and adapt to their
shortcomings. Basic research is required now to advance methods in
time for these new machines to become operational. As machines begin
to operate at hundreds of petaflops peak performance and beyond,
trends toward reduced energy consumption will require incredibly high
levels of concurrency to achieve the desired computation rates
\cite{kogge_using_2011}. The end result of these hardware changes is
that the larger number of low-powered processors will be prone to both
soft failures such as bit errors in floating point operations and hard
failures where data cannot be recovered. Because these failures are
predicted to be common, resilient solver technologies are required to
overcome these events at the application level. With linear solvers
based on Monte Carlo techniques, such issues are potentially
alleviated by statistical arguments. In the case of soft failures,
isolated floating point errors in the Monte Carlo simulation are
absorbed within tally statistics. Completely losing memory during a
hard failure is treated as a high variance event where some portion of
the Monte Carlo histories and subsequently the solution are lost with
other histories maintained by replicating the problem and combining
the results through superposition.

In this paper, we will present the Monte Carlo methods of Neumann and
Ulam and then outline Monte Carlo Synthetic Acceleration methods. We
then present a parallel algorithm for MCSA based on the work of
Brunner and Brantley \cite{brunner_efficient_2009} and Wagner
et. al. \cite{wagner_hybrid_2010}. Using this algorithm, we perform a
series of scaling studies to characterize parallel
performance. Finally, we summarize the results and offer future work
that may improve the algorithm.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Monte Carlo Methods}
Monte Carlo Synthetic Acceleration methods rely on the Neumann-Ulam
class of linear solvers to accelerate a stationary method. These Monte
Carlo methods work by sampling a distribution with an expectation
value equivalent to that of the inverted operator. Neumann and Ulam's
work on these solvers was first published in 1950 by Forsythe and
Leibler \cite{forsythe_matrix_1950}. In the following decade, the
books of Hammersley and Handscomb \cite{hammersley_monte_1964} and
Spanier and Gelbard \cite{spanier_monte_1969} present additional
detail on the methods. In this section we briefly present the
Neumann-Ulam Monte Carlo method for solving linear systems and then
present Monte Carlo Synthetic Acceleration. We refer the reader to
\cite{evans_monte_2009} and \cite{evans_monte_2012} for a more
detailed explanation and analysis of MCSA and the Neumann-Ulam
methods.

\subsection{Neumann-Ulam Method}
We seek solutions of the general linear problem in the following form:
\begin{equation}
  \mathbf{A} \mathbf{x} = \mathbf{b}\:,
  \label{eq:linear_problem}
\end{equation}
where $\mathbf{A} \in \mathbb{R}^{N \times N}$ is a matrix operator
such that $\mathbf{A} : \mathbb{R}^{N} \rightarrow \mathbb{R}^{N}$,
$\mathbf{x} \in \mathbb{R}^N$ is the solution vector, and $\mathbf{b}
\in \mathbb{R}^N$ is the forcing term.  For a given linear operator
$\mathbf{A}$, we can use diagonal splitting to define the
\textit{iteration matrix}, $\mathbf{H}$:
\begin{equation}
  \mathbf{H} = \mathbf{I} - \mathbf{A}\:,
  \label{eq:linear_mc_iteration_matrix}
\end{equation}
such that we are solving the system:
\begin{equation}
  \mathbf{x} = \mathbf{H} \mathbf{x} + \mathbf{b}\:.
  \label{eq:richardson_split}
\end{equation}
We can then form an alternative representation for $\mathbf{A}^{-1}$
by generating the \textit{Neumann series} solution:
\begin{equation}
  \mathbf{A}^{-1}\mathbf{b} = \sum_{k=0}^{\infty}
  \mathbf{H}^k\mathbf{b} = \mathbf{x}\:,
  \label{eq:neumann_solution}
\end{equation}
which will converge if the spectral radius of $\mathbf{H}$ is less
than 1. If we expand the summation with a succession of matrix-vector
multiply operations, we arrive at an alternative perspective of this
summation by considering the $i^{th}$ component of the solution
vector:
\begin{equation}
  x_i = \sum_{k=0}^{\infty}\sum_{i_1}^{N}\sum_{i_2}^{N}\ldots
  \sum_{i_k}^{N}h_{i,i_1}h_{i_1,i_2}\ldots h_{i_{k-1},i_k}b_{i_k}\:,
  \label{eq:expanded_neumann_solution}
\end{equation}
which can interpreted as a series of transitions between states,
\begin{equation}
 \nu = i \rightarrow i_1 \rightarrow \cdots \rightarrow i_{k-1}
 \rightarrow i_{k}\:,
  \label{eq:mc_walk_permutation}
\end{equation}
in $\mathbf{H}$ where $\nu$ is interpreted as a particular sequence
permutation. We can generate these sequences of transitions through
Monte Carlo random walks by assigning them both a probability and
weight. Using the adjoint form \cite{spanier_monte_1969},
probabilities are column-scaled:
\begin{equation}
  p_{ij} = \frac{|h_{ji}|}{\sum_j |h_{ji}|}\:,
  \label{eq:adjoint_probability}
\end{equation}
such that we expect to select a new state, $j$, from the current state
in the random walk, $i$, by sampling the probability distribution
function formed by the $i^{th}$ row of the probability matrix. The
transition weight is then defined as:
\begin{equation}
  w_{ij} = \frac{h_{ji}}{p_{ij}}\:.
  \label{eq:adjoint_weight}
\end{equation}
The initial state $i_0$ of the random walk is determined by sampling
the source vector $\mathbf{b}$ with initial probabilities and weights:
\begin{equation}
  P_{(i_0=i)}(\nu) = \frac{|b_i|}{||\mathbf{b}||_1},\ \ \ W_0 =
  ||\mathbf{b}||_1 \frac{b_i}{|b_i|}\:.
  \label{eq:adjoint_starting_params}
\end{equation}
The contribution to the solution in state $j$ from a particular random
walk permutation of $k$ events is then the \textit{collision
  estimator}:
\begin{equation}
  X_{j}(\nu) = \sum_{m=0}^k W_{m} \delta_{i_m,j}\:,
  \label{eq:adjoint_permutation_contribution}
\end{equation}
where the Kronecker delta indicates that the tally contributes only in
the current state, $i_m$, of the random walk and 
\begin{equation}
  W_{m} = W_0 w_{i_0,i_1} w_{i_1,i_2} \cdots w_{i_{m-1},i_m}\:.
  \label{eq:adjoint_permutation_weight}
\end{equation}
Finally, the expectation value using all permutations is:
\begin{equation}
  E\{X_j\} = \sum_{\nu} P_{\nu} X_{j}(\nu)\:,
  \label{eq:adjoint_expectation_value}
\end{equation}
with
\begin{equation}
  P_{\nu} = P_{(i_0=i)} p_{i,i_1} p_{i_1,i_2} \cdots p_{i_{k-1},i_k}\:.
  \label{eq:adjoint_permutation_probability}
\end{equation}
Finally, we set a criteria for random walk termination by setting a
\textit{relative weight cutoff}:
\begin{equation}
  W_f = W_c W_0\:,
  \label{eq:relative_weight_cutoff}
\end{equation}
where $W_c$ is a user defined cutoff. The random walk will then be
terminated after $m$ steps if $W_m < W_f$.

\subsection{Monte Carlo Synthetic Acceleration}
Using the ideas of Halton's residual method
\cite{halton_sequential_1994}, Evans and Mosher recently developed a
Monte Carlo solution method that was not prohibited severely by the
quality of the initial guess for the system \cite{evans_monte_2009}
and later applied it more rigorously as a solution mechanism for the
radiation diffusion equation \cite{evans_monte_2012}. With their new
methods, they achieved identical numerical results when compared to
conventional subspace methods and observed better performance with
respect to time to solution. Their approach was instead to use
residual Monte Carlo as a synthetic acceleration for a stationary
method.

The \textit{Fixed-Point Monte Carlo Synthetic Acceleration} method is
defined as:
\begin{subequations}
  \begin{gather}
    \label{eq:mcsa_1}
    \mathbf{r}^{k} = \mathbf{b} - \mathbf{A}\mathbf{x}^{k}\:,\\
    \label{eq:mcsa_2}
    \mathbf{x}^{k+1/2} = \mathbf{x}^k + \mathbf{r}^k\:,\\
    \label{eq:mcsa_3}
    \mathbf{r}^{k+1/2} = \mathbf{b} - \mathbf{A}\mathbf{x}^{k+1/2}\:,\\
    \label{eq:mcsa_4}
    \mathbf{A}\delta\mathbf{x}^{k+1/2} = \mathbf{r}^{k+1/2}\:,\\
    \label{eq:mcsa_5}
    \mathbf{x}^{k+1} = \mathbf{x}^{k+1/2} + \delta \mathbf{x}^{k+1/2}\:,
  \end{gather}
  \label{eq:mcsa}
\end{subequations}
where $\mathbf{r}$ is the residual vector, $\delta\mathbf{x}$ is the
Monte Carlo correction vector, and $k$ indicates the iteration
index. Here, a Neumann-Ulam Monte Carlo method is used to generate the
solution correction in Eq~(\ref{eq:mcsa_4}) using the residual as
source and Richardson's iteration in Eq~(\ref{eq:mcsa_2}) has been
rewritten as a residual correction. Using Monte Carlo in this way
achieves the same effect as Halton's method, decoupling its
convergence rate from the overall convergence rate of the method. For
the Neumann-Ulam method, a set number of random walk permutations is
specified to generate the correction at each MCSA iteration. Adding
more histories to the Neumann-Ulam sequence reduces the statistical
uncertainty in $\delta\mathbf{x}$ and further accelerates the
fixed-point iteration. In addition to the Monte Carlo solver
parameters dictating the number of histories and weight cutoff, the
outer MCSA iterations also have the following stopping criteria:
\begin{equation}
  ||\mathbf{r}||_\infty < \epsilon \ ||\mathbf{b}||_\infty\:,
  \label{eq:mcsa_stopping_criteria}
\end{equation}
where $\epsilon$ is a user-defined parameter.

@@

\subsection{Overlapping Domain Decomposition}
\label{subsec:overlapping_domain_decomp}
In the next set of scaling studies, we will explore how adding overlap
to the subdomains in the problem affects run times and parallel
performance as compared to the case with pure domain decomposition. To
do this, we utilize the set of analytic relations developed in
\cite{slattery_spectral_2013} to design these numerical
experiments. For the neutron diffusion problem used in these scaling
studies, the linear operator had a spectral radius of 0.787 and a
weight cutoff of \sn{1}{-2} was used with the Neumann-Ulam
solver. Using Eq~(\ref{eq:analytic_k}) gives an expected random walk
length of 19.22 transitions for any given history in the
problem. Using the mean-chord approximation given by
Eq~(\ref{eq:mean_chord_domain_leakage}), we then expect approximately
5.3\% of the histories born in a particular domain to leak out after
the first transport stage. By adding overlap, we aim to reduce this
fraction of communicated histories and therefore boost the parallel
performance of the algorithm.
